---
title: "NYPD Shooting Incident Data Report"
author: "Y.Saito"
date: '2023-03-03'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importance of reproducibility

Reproducibility is important factor to provide objectivity, persuasiveness with its underlying data, and future research/analysis.

### The steps in the data science process

The key elements of the data science are;

1. Explore
2. Wrangle
3. Program
4. Model
5. Communicate

source: The table of contents from a book "**R for Data Science**" by Hadley Wickham and Garrett Grolemund 
https://r4ds.had.co.nz/

Here are the steps on this report;

- Search relevant source data
- Import data
- Look at data
    - *and tidy*
- Visualize data 
- Output data in report format

In a column 'PERP_AGE_GROUP', some error figures were found, such as {224,940,1020}. They should be replaced with 'UNKNOWN'.

For the columns 'PERP_AGE_GROUP',	'PERP_SEX',	and 'PERP_RACE', about 40% of records were either missing or 'UNKNOWN'. Though the remaining 60% of valid records could produce some valuable information.

```{r}
# Import libraries
library(tidyverse)
library(lubridate)
```

## Importing data

```{r}
# Import external csv data 
url_in <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
cases <- read_csv(url_in)
```

```{r summary of NYPD shooting incidents data}
# check data
glimpse(cases)
summary(cases)
```
```{r Month}
# Change the data type from "character" to "date"
dates <- cases %>%
  mutate(OCCUR_DATE = lubridate::mdy(OCCUR_DATE))
# Extract number of incidents each month
m_table <- table(month(dates$OCCUR_DATE))
# Plot
barplot(m_table, main="Monthly Trend")
```

It shows the monthly trend. In winter, the number of incidents goes down.
```{r Hour}
# Change the data type from "character" to "period"
times <- cases %>%
  mutate(OCCUR_TIME = lubridate::hms(OCCUR_TIME))
# Extract number of incidents each hour
h_table <- table(hour(times$OCCUR_TIME))
# Plot
plot(h_table, main="Hourly Trend", col=rgb(1,0,1), type="l")
```

The incident rate is the highest around the midnight, and the lowest in a morning, around commuting hours. 

```{r location}
# Generate a table; List of Boroughs(rows) and Number of incidents(column).
# Prepare empty table
loca_table <- matrix(0,5,3)
# Set column names
colnames(loca_table) <- c("incidents","population","per 1000")
# Set row names
rownames(loca_table) <- rownames(table(cases$BORO))
# Insert the number of incidents in each Boroughs
loca_table[,1] <- table(cases$BORO)
# Insert the population in each Boroughs 
loca_table[,2] <- c(1472654,2736074,1694251,2405464,495747)
# Insert the incidents-rate per 1000 people
loca_table[,3] <- loca_table[,1] / loca_table[,2] * 1000
# Output the table
loca_table
```
Boroughs population data from;
http://www.citypopulation.de/en/usa/newyorkcity/
(referred date: 2023/03/03)
```{r pi-chart}
# Insert BORO-wise incidents-rate per 1000 into a variable
pie.per1000 <- c(loca_table[,3])
# Calculate percents
pct <- round(c(loca_table[,3])/sum(c(loca_table[,3]))*100) 
# Add percents to labels
lbls <- paste(rownames(loca_table), pct) 
# Add % to labels
lbls <- paste(lbls,"%",sep="") 
# Plot
pie(pie.per1000,labels = lbls, main="Location Trend")

```

Location Trend is based on Incident Rate per 1000 people (= Number of incidents / Population * 1000); 
Further analysis should be held on age groups, races, locations + races, locations + age groups, and so forth.
```{r Murder Flag,include=FALSE}
# 要素数をMFlg別にカウントして円グラフ作成
MurderFlg <- cases %>% select(STATISTICAL_MURDER_FLAG
)
MurderFlg  
```
```{r Victim Age Group,include=FALSE}
# 要素数を被害者年齢別にカウントして棒グラフ作成
VicAge <- cases %>% select(VIC_AGE_GROUP)
VicAge  
```

## Conclusion and Bias
The data is cumulative over the period of 2006-2021. The analysis of “Monthly Trend”, “Daily Trend”, and “Location Trend” over the period is worth comparing to the latest data, i.e., an analysis with the data of just 2021. By comparing them yearly basis, we may find how the urban development progress, as well as the economy, affects the incidents.

An example of bias the data may consist of is the column ‘PERP_RACE’. The number of a certain minority race is extremely high (10668/25596) compare to the other races(the sum of five other races was just 3782/25596, while the data also contains ‘NA’ and ‘UNKNOWN’ which sum is 11146/25596. There could be a some sort of manipulation or intention over these figures. My personal bias is as well the observation above. I see the US as a nation of diversity, thus I naturally have paid attention to the ‘PERP_RACE’ column, but this is just my bias. I must pay attention to other columns and any correlations to reduce the effect of a personal bias over the analysis.